CF Registry Analysis

### 2024-03-13 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
remove ABPA from survival models

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    fig-width: 10
    fig-height: 10
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
execute:
  echo: false
  warning: false
  cache: true
editor: source
---

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_pulmonarycomplications1", "A_corticosteroids1",
  "cfrd_status_annual", "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID", birth_year = "Birth year", tstart = "Interval start", tstop = "Interval stop", age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
# Labels
label(s) <- labels[colnames(s)]
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

>>>>

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    fig-width: 10
    fig-height: 10
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual", 
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID", birth_year = "Birth year", tstart = "Interval start", tstop = "Interval stop", age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
# Labels
label(s) <- labels[colnames(s)]
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation",  "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```
